---
title: "After the first round"
author: "Tom Kain"
date: "January 6, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(DT)
library(magrittr)
library(googlesheets)
library(tidyverse)
```

First four games in the books and I'm already out. Let's see where things lie.

```{r import, message=FALSE}
raw_df <- read_csv('outcomes_analysis1.csv') %>%
  filter(
    `AFC Wildcard 1` == "Chargers" &
    `AFC Wildcard 2` == "Colts" &
    `NFC Wildcard 1` == "Cowboys" &
    `NFC Wildcard 2` == "Eagles"
    ) %>%
  select(-(`AFC Wildcard 1`:`AFC Wildcard 2`), 
         -(`NFC Wildcard 1`:`NFC Wildcard 2`))

raw_df %>%
  summarize(p = sum(p)) %>%
  pull() -> p_sum

init_df <- raw_df %>%
  mutate(p_adj = p * 1/p_sum)

```

```{r initreview}
outcome_count <- init_df %>% nrow()

init_df %>% 
  group_by(Winner) %>%
  summarize(n = n(),
            perc = n/outcome_count,
            p_adj = sum(p_adj)) %>%
  arrange(desc(n))
```

Looks like we're down to `r outcome_count` possibilities left. Dad had the biggest net gain, nearly tripling his share of the available outcomes. Will drops a little less than 10 percent.

What are the scenarios in which a tie could occur? 


```{r review2}
init_df %>%
  filter(Winner == 'Tie') %>%
  select(-(Dad_total:Will)) %>%
  datatable()
```

If the Eagles, Chargers, or Colts win again, there can be no tie. Otherwise, there's an outside chance for a tie.

Who is Dad rooting for in round 2? 

```{r review3}
init_df %>%
  filter(Winner == 'Dad') %>%
  select(-(Dad_total:Will)) %>%
  datatable()
```

Dad needs the following all to happen:

1. Pats beat the Chargers
2. Saints beat the Eagles
3. Cowboys beat the Rams
4. One of his teams (Pats, Saints, Cowboys) win the Super Bowl